\newpage
\begin{problem}{20}{2}{2}
  Show that as the stiffness~$\beta$ goes to~$\infty$, the soft K-means algorithm becomes identical to the original hard K-means algorithm except for the way in which means assigned no points behave.  Describe what those means do instead of sitting still.
\end{problem}

In the standard or ``hard'' K-means algorithm, each point is assigned to exactly one cluster.  As such, each of a cluster's points have equal membership.

Soft K-means reduces the rigidity of the standard K-means by introducing a new ``stiffness'' hyperparameter,~$\beta$.  Rather than each point being a member of exclusively one cluster, the \textit{responsibility} for each point is shared (generally unevenly) among all $K$~clusters.  For cluster~$k\in\{1,\ldots,K\}$ and point~$\textbf{x}^{(n)}$, the responsibility,~$r_{k}^{(n)}$, is:

\[ r_{k}^{(n)} = \frac{\exp(-\beta d(\textbf{m}^{k}, \textbf{x}^{(n)}))}{\sum_{k'} \exp(-\beta d(\textbf{m}^{k'}, \textbf{x}^{(n)}))} \]

\noindent
where $d$ is the distance metric, and $\textbf{m}^{k}$ is the center of cluster~$k$.

As $\beta$ increases, then even small differences in $d$ can cause massive changes in responsibility.  As ${\beta \rightarrow \infty}$, all responsibility for a point will be assigned to its nearest cluster.  This behavior is exactly the same as standard, ``hard'' K-means where points belong to only the cluster's whose centroid is closest.

Considering the second part of the question, the centroid update rule for cluster,~$k$, is:

\[ \mathbf{m}^{(k)} = \frac{\sum_{n}r^{(n)}_k \mathbf{x}^{(n)}}{R^{(k)}} \]

\noindent
where the total responsibility, $R^{(k)}$, for cluster $k$ is:

\[ R^{(k)}=\sum_{n} r_{k}^{(n)} \text{.} \]

As mentioned previously, small differences in the distance,~$d$, cause huge differences in the responsibility,~$r$, when $\beta$ approaches infinity.  In hard K-means, it was possible for a cluster to be assigned no points.  However, in soft K-Means, such clusters still have \textit{some} responsibility for every point, albeit infinitesimally small when $\beta$ approaches infinity.  In that case, those means will move to the \textit{location of the point with the highest responsibility, which may not necessarily be the closest to that mean}.
